\chapter{Implementation}
\label{chap:implementation}
This chapter describes the practical implementation of the \acrlong{ew} system, building upon the designs introduced in the previous chapters. While \cref{chap:onchainDesign} and \cref{chap:offchainDesign} focus on the system's architecture and the rationale behind key design decisions, this chapter highlights the concrete tools, technologies, and development processes used to bring the system to life.

Key implementations aspects include the use of account abstraction for users account, the integration of off-chain and on-chain components and the introduction of an access control system that empowers students to manage who can access their data. Special attention is given to the coordination between the academic records system and the decentralized storage layer. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DEVELOPMENT ENVIRONMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Development Environment}
This section outlines the tools and technologies used to author, compile, and test \acrlong{ew} system.

The primary code editor is Visual Studio Code\footnote{\url{https://code.visualstudio.com/}}, chosen for its lightweight footprint and extensive ecosystem of extensions. To support Solidity development, we installed the Nomic Foundation's Solidity extension, which provides syntax highlighting, inline error detection, and code completion directly within the editor, thereby accelerating development and debugging.

This extension provides syntax highlighting, inline error detection, and code completion directly within the editor, thereby accelerating development and debugging.

To compile contracts and simulate a local blockchain, we utilize Hardhat\footnote{\url{https://hardhat.org/}}. Our configuration specifies:
\begin{itemize}
    \item \textbf{Solidity compiler version}: 0.8.28 (the latest fully supported by Hardhat at the time of writing)\footnote{Versions 0.8.29 and 0.8.30 were not yet fully supported.}
    \item \textbf{EVM version}: Cancun
    \item \textbf{Optimizer settings}: Enabled with 1000 runs to reduce bytecode size and keep contracts (e.g., \texttt{Student}, \texttt{EntryPoint}) below the 24576-byte limit.
    \item \textbf{Preconfigured account}: A single funded \acrshort{eoa} serves as deployer, funder for student and university wallets, and system administrator. Its private key is fixed to ensure deterministic contract addresses, as these are derived from the deployer’s key and the number of contracts deployed by the account (\gls{nonce}).
\end{itemize}
The complete Hardhat configuration file can be found in \cref{chap:hardhatConf}.

For blockchain interaction in off-chain code, we rely on two core libraries:
\begin{itemize}
\item \textbf{ethers}: Provides providers for network connectivity, cryptographic utilities (e.g., PBKDF2), and the \texttt{Wallet} class for \acrshort{eoa} key management and transaction signing.
\item \textbf{TypeChain}: A Hardhat plug-in that generates TypeScript bindings from contract \acrfull{abi}, enabling contracts to be imported and used as strongly typed classes.
\end{itemize}

Finally, our smart contracts leverage external libraries to streamline common patterns:
\begin{itemize}
\item \textbf{OpenZeppelin Contracts}: Supplies contract implementations of access control and other standard modules.
\item \textbf{AccountAbstraction}: Provides the EntryPoint contract and abstract contracts required for ERC-4337 account abstraction.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ON-CHAIN INTEGRATION IN OFF-CHAIN COMPONENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Account Abstraction and On-Chain Integration in Off-Chain Components}
As described in \cref{ssec:accountAbstraction}, account abstraction redefines on-chain interactions within the \acrlong{eth} network. Whereas traditions transactions were bound to \acrlong{eoa}, the introduction of UserOperations enables \acrlong{sca} to initiate and execute transactions directly on-chain. This paradigm shift also alters how off-chain components interact with and invoke on-chain logic.

In \cref{ssec:extBlockchainInteraction}, we identified three primary modalities for invoking smart contract functionality:
\begin{enumerate}
    \item Direct view function calls
    \item View function calls via a \acrlong{sca}
    \item Gas-consuming transactions via a \acrlong{sca}
\end{enumerate}
The following subsections detail the code patterns used to implement each interaction modality, as well as the core components involved in the implementation of account abstraction.

\subsection{Direct View Function Calls}
To perform a direct call to the view function shown in \cref{lst:directFuncCall}, the system first initializes an ethers provider. It then uses the StudentsRegister\_\_factory class generated by TypeChain to create a StudentRegister instance connected to the on-chain contract via the \textit{connect} method, which accepts the contract address and provider as arguments. The instance exposes all public functions, variables, and types defined in the Solidity contract. 
Because the target function is permissioned, the student's \acrshort{eoa} must be connected as the transaction sender. This is achieved by invoking the factory's \textit{connect} method again, this time passing a \textit{Wallet}, instantiated with the student's private key and the same provider. The view function is then called as a standard method on the resulting object. The return value is automatically typed according to the contract's Solidity definition and, in this case, is a string.

% TODO: Talk about advantages of this interaction system

\lstinputlisting[
    caption={Direct call to a smart contract view function},
    label=lst:directFuncCall,
    language=TypeScript
]{listings/directFunctionCall.ts}

\subsection{View Function Calls via a Smart Contract Account}
\label{ssec:viewFunctionCalls}
View functions calls via a \acrshort{sca} exploit the \textit{executeViewCall} method defined in the SmartAccount abstract contract (see \cref{lst:executeViewCall}). This method, which can be only invoked by the \acrshort{sca} owner, accepts two arguments: the address of the target contract and encoded function name and parameters. By using it, \acrshort{sca}s can perform read-only operations without submitting gas-consuming transactions. 
\lstinputlisting[
    caption={\textit{executeViewCall} method in the \textit{SmartAccount} abstract contract.},
    label=lst:executeViewCall,
    language=Solidity
]{listings/executeViewCall.sol}

To invoke this method from TypeScript, the steps illustrated in \cref{lst:executeViewCallTS} must be followed:
\begin{enumerate}
    \item \textbf{Calldata encoding}: The target function’s selector and parameters must be encoded into a hexadecimal string. We use the \textit{BaseContract} class from \textit{ethers} to generate this encoding in a generic way, independent of the specific contract interface.
    \item \textbf{Account connection and method invocation}: Instantiate a \textit{SmartAccount} contract object connected with the owner’s \acrshort{eoa} wallet (e.g., the university’s \acrshort{eoa}). Then, invoke the \textit{executeViewCall} method on this connected \acrshort{sca} instance, passing the target contracts address along with the encoded calldata.
    \item \textbf{Result decoding}: The call returns an encoded hexadecimal string. We decode it back into expected return type using \textit{ethers} utilities.
\end{enumerate}
\lstinputlisting[
    caption={TypeScript code invoking \textit{executeViewCall} on a \textit{SmartAccount} instance.},
    label=lst:executeViewCallTS,
    language=TypeScript
]{listings/executeViewCall.ts}

\subsection{Gas-Consuming Transactions via Smart Contract Account}
To execute state-changing operations, such as granting or revoking permissions, our off-chain components fully leverage the ERC-4337 account abstraction protocol. On the smart-contract side, the \textit{BaseAccount} abstract contract provides two core methods (see \cref{chap:baseAccount}):
\begin{enumerate}
    \item \textbf{execute}: Accepts a target address (either an \acrshort{eoa} or another smart contract), a value (\acrshort{eth} to transfer) and encoded calldata, then performs a single transaction.
    \item \textbf{executeBatch}: Accepts an array of such triples (address, value, calldata) and executes them automatically in a single transaction.
\end{enumerate}
Both methods incur in gas costs, in contrast to the read-only \textit{executeViewCall} from \cref{ssec:viewFunctionCalls}.

Account abstraction security is enforced by the \textit{validateUserOp} function within BaseAccount, invoked by the EntryPoint contract before execution. Within \textit{validateUserOp}, our smart account calls its own \textit{\_validateSignature} helper (see \cref{lst:validateSignature}),  which in our case simply verifies that the UserOperation’s signer matches the account owner.
\lstinputlisting[
    caption={Function to validate the sender's signature in the \textit{SmartAccount} contract.},
    label=lst:validateSignature,
    language=Solidity
]{listings/validateSignature.sol}

On the client side, we encapsulate UserOperation handling in an \textit{AccountAbstraction} TypeScript class (see \cref{chap:accountAbstraction}). Because no public bundlers are available in our local testnet, we must format and submit operations directly to the EntryPoint contract as \textit{PackedUserOperation} objects. The workflow is as follows:
\begin{enumerate}
    \item \textbf{Create the UserOperation}: Populate fields such as sender address, target contract address, value, calldata, gas and fee parameters, and paymaster details (see \cref{ssec:accountAbstraction}).
    \item \textbf{Pack the UserOperation}: pack and encode some UserOperation fields, such as the gas consumption ones.
    \item \textbf{Sign the UserOperation}: Sign the packed payload with the sender's \acrshort{eoa} using \textit{ethers} utilities, incorporating some network information and operation format.
    \item \textbf{Send the UserOperation}: Submit the signed operation via the \textit{EntryPoint.handleOps} method. Because this submission itself is a transaction, the \acrshort{cli} must pre-fund users' \acrlong{eoa}s on the testnet. 
    \item \textbf{Verify the result}: After execution, inspect the returned execution trace stack for any exception logs emitted by the UserOperation. An absence of such errors confirms successful execution.
\end{enumerate}

The primary limitation of our implementation lies in gas consumption estimation. In our AccountAbstraction class, all gas and fee-related parameters are hardcoded. Since the system operates in a controlled testing environment with dedicated paymaster and user accounts pre-funded with large balances, we configured these limits conservatively, setting them to high values to avoid transaction failures due to out-of-gas errors.
We chose not to implement a more advanced gas estimation mechanism, as it would introduce unnecessary complexity in the context of a local setup. In a real-world deployment, this concern is mitigated by bundlers, which either expose utilities for gas estimation or handle it internally. As a result, the interaction with UserOperations becomes seamless for the end user and developers alike.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACCESS CONTROL SYSTEM
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Access Control System}
One of the main principles of Web3 is the shift in data ownership and management, from centralized systems typical of Web2 architectures to a user-centric model. In \acrlong{ew}, academic records are owned by students, as established by \textit{FR 9} in \cref{tab:funcReq}, and this onwership is enforced through a robust access control system. This permission system is implemented entirely in smart contracts, specifically within the Student contracts.

Access control in the Student contract is achieved using the \textit{AccessControlEnumerable} abstract contract, an extension of the more general \textit{AccessControl} contract, both of which are provided by the OpenZeppelin contracts library. This library allows developers to define roles, which are then used to restrict access to specific contract functions. In AccessControlEnumerable, a role is defined as the \Gls{keccak256} \gls{hash} of a string, as shown in \cref{lst:roleDefinition}.
\lstinputlisting[
    caption={Role definition in the \textit{Student} smart contract},
    label=lst:roleDefinition,
    language=Solidity
]{listings/roleDefinition.sol}

Roles can be granted or revoked using methods provided by the AccessControlEnumerable library. The \textit{Student} contract utilizes the following methods:
\begin{itemize}
    \item grantRole
    \item revokeRole
    \item \_grantRole
    \item \_revokeRole
\end{itemize}
The key distinction between these methods is their access control. The functions \textit{grantRole} and \textit{revokeRole} can only be invoked in a transactions initiated by the entity holding the DEFAULT\_ADMIN\_ROLE, whereas \textit{\_grantRole} and \textit{\_revokeRole} are unrestricted methods. The DEFAULT\_ADMIN\_ROLE is defined by the AccessControlEnumerable library and must be assigned to the administrator of the roles, in our case the student. This role is granted at the time of smart account creation. The unrestricted methods are primarily used in functions involving universities (e.g., permission requests), while the permissioned methods are used by students (e.g., permission revocation).

To enforce role-based access, the contract uses the \textit{onlyRole} function modifier (\cref{lst:roleModifier}) and the \textit{hasRole} method (\cref{lst:roleMethod}). The modifier restricts access to function execution based on role possession, while the method allows for role verification at runtime. Both tools are provided by the OpenZeppelin access control library.
\lstinputlisting[
    caption={Function definition using the \texttt{onlyRole} modifier.},
    label=lst:roleModifier,
    language=Solidity
]{listings/rolesFunctionModifier.sol}
\lstinputlisting[
    caption={\textit{hasRole} method for verifying access permissions.},
    label=lst:roleMethod,
    language=Solidity
]{listings/rolesMethod.sol}

As shown in \cref{lst:askAndGrantPermission}, the permission-related functions accept or return roles in their \gls{hash}ed format. Therefore, off-chain components must define and compute permissions in the same way as the contract.

\lstinputlisting[
    caption={Functions to request and grant permissions in the \textit{Student} contract.},
    label=lst:askAndGrantPermission,
    language=Solidity
]{listings/askAndGrantPermission.sol}
Both the browser extension and the \acrshort{sdk} use the \textit{id} function from the \textit{ethers} library to compute the \gls{hash} of role strings, as shown in \cref{lst:offchainRole}. The \textit{id} function takes a UTF-8 string and returns its \Gls{keccak256} \gls{hash}, ensuring consistency with the contract's role definition.   
\lstinputlisting[
    caption={Role \gls{hash} generation in off-chain components.},
    label=lst:offchainRole,
    language=TypeScript
]{listings/offchainRoleDefinition.sol}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DECENTRALIZED STORAGE SYSTEM INTERACTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Decentralized Storage System Interaction}
\acrlong{ew} employs a decentralized storage system to manage certificates linked to academic records. Both the browser extension and the \acrshort{sdk} can retrieve files, but only the \acrshort{sdk} is responsible for uploading them on it.

File retrieval is performed via the public \acrshort{ipfs} gateway \acrshort{url}. By appending a file's \acrshort{cid} to this gateway \acrshort{url}, students and universities can directly access the document. This gateway is hardcoded in both the browser extension and the \acrshort{sdk}, as shown in \cref{lst:ipfsAccess}. The corresponding \acrshort{cid} for each certificate is stored on-chain and retrieved when querying academic results.
\lstinputlisting[
    caption={\acrshort{ipfs} public gateway \acrshort{url} configuration in off-chain components.},
    label=lst:ipfsAccess,
    language=TypeScript
]{listings/ipfsAccessConf.ts}

Uploading a file requires interaction with an \acrshort{ipfs} pinning service, as described in \cref{ssec:pinningService}. Our \acrshort{sdk} uses an \acrshort{aws} S3–compatible \acrshort{api} provided by Filebase (see \cref{chap:pinningAPI}). The API call requires the following parameters:
\begin{itemize}
    \item \textbf{Bucket}: Name of the virtual folder for organizing files.
    \item \textbf{Key}: Unique identifier for the file within the bucket.
    \item \textbf{Body}: The file content to be uploaded.
    \item \textbf{ContentType}: The type of the file (set to \acrshort{pdf} in our implementation).
\end{itemize}
Prior to issuing the upload request, a middleware, sourced from the official Filebase documentation\footnote{\url{https://docs.filebase.com/}}, is attached to the API client. This middleware intercepts the response to extract the resulting \acrshort{cid}, which the SDK then records on-chain alongside the academic record.
